Faithful Variable Screening for High-Dimensional Convex Regression

### 2014-07-22 tex/scam/estimation.tex
redoing section as subsection of additive_faithful, expanding on algo description

\section{Estimation Procedure}
\label{sec:acdc}

Theorem~\ref{thm:acdc_faithful} suggest a two stage procedure for variable selection in convex regression--first learning a sparse convex additive model and then, on the residual, separately learning an univariate concave function for each of the dimensions.

More precisely, given samples $(\mathbf{x}_1, y_1), ..., (\mathbf{x}_n, y_n)$, we perform the following procedure, which we refer to as AC/DC (additively convex / decoupled concave):
\begin{enumerate}
\item Compute: 
\begin{equation}
\label{eqn:scam}
\hat{f}_1, ..., \hat{f}_p = \argmin_{f_1,...,f_p \in \mathcal{C}^1} 
   \frac{1}{n} \sum_{i=1}^n \Big(y_i - \sum_{k=1}^p f_k(x_{ki}) \Big)^2 
       + \lambda \sum_{k=1}^p \| f_k \|_\infty
\end{equation}
\item Compute, for each $k$ such that $\| \hat{f}_k \| = 0$:
\begin{equation}
\label{eqn:dc}
\hat{g}_k = \argmin_{g_k \in \mathcal{C}^1} 
   \frac{1}{n} \sum_{i=1}^n \Big( y_i - \sum_{k'} \hat{f}_{k'}(x_{k'i}) 
    - g_k(x_{ki})\Big)^2 
      + \lambda \| g_k \|_\infty
\end{equation}
\item Output as the set of relevant variables: 
$\{ k \,:\, \| \hat{f}_k \|_\infty > 0 
  \textrm{ or } \| \hat{g}_k \|_\infty > 0 \}$ 
\end{enumerate}

We added an $\ell_\infty/\ell_1$ penalty in equation~\ref{eqn:scam} and the $\ell_\infty$ penalty in equation~\ref{eqn:dc} to encourage sparsisty. There are other forms of penalty that can also produce sparse estimates such as a penalty on the derivative of each of the component functions; we find the form we use to be convenient both for theoretical analysis and implementation.

>>>>

\subsection{Estimation Procedure}
\label{sec:acdc}

Theorem~\ref{thm:acdc_faithful} naturally suggests a two-stage screening procedure for
variable selection.  In the first stage we fit a sparse convex additive model.
In the second stage, for every variable marked as irrelevant in the first stage,
we fit a univariate \emph{concave} function separately on the residual for
that variable. We refer to this procedure as AC/DC (additive
convex/decoupled concave).

More precisely, given samples  
$(\mathbf{x}_1, y_1), ..., (\mathbf{x}_n, y_n)$, 
we perform the following procedure, which we refer to as AC/DC (additively convex / decoupled concave):
\begin{enumerate}
\item {\it AC Stage}: Estimate an additive convex model
\begin{equation}
\label{eqn:scam}
\hat{f}_1, ..., \hat{f}_p, \hat \mu  = \argmin_{f_1,...,f_p \in
  \mathcal{C}^1, \mu\in\reals} 
   \frac{1}{n} \sum_{i=1}^n \Big(y_i - \mu - \sum_{k=1}^p f_k(x_{ik}) \Big)^2 
       + \lambda \sum_{k=1}^p \| f_k \|_\infty.
\end{equation}
\item {\it DC Stage}: For each $k$ such that $\| \hat{f}_k \| = 0$, estimate
  a concave function on the residual:
\begin{equation}
\label{eqn:dc}
\hat{g}_k = \argmin_{g_k \in \mathcal{C}^1} 
   \frac{1}{n} \sum_{i=1}^n \Big( y_i - \hat \mu - \sum_{k'} \hat{f}_{k'}(x_{ik'}) 
    - g_k(x_{ik})\Big)^2 
      + \lambda \| g_k \|_\infty.
\end{equation}
\item Output as the set of relevant variables
$\hat S = \{ k \,:\, \| \hat{f}_k \|_\infty > 0 
  \textrm{ or } \| \hat{g}_k \|_\infty > 0 \}$. 
\end{enumerate}

We use an $\ell_\infty/\ell_1$ penalty in equation~\eqref{eqn:scam}
and an $\ell_\infty$ penalty in equation~\eqref{eqn:dc} to encourage
sparsity.  Other penalties can also produce
sparse estimates, such as a penalty on the derivative of each of the
component functions.  The $\|\cdot\|_\infty$ norm convenient for both
theoretical analysis and implementation.

The optimization in 
\begin{align}
\label{eq:convreg}
\begin{split}
\text{minimize}_{f, \beta} & \;\; \sum_{i=1}^n (Y_i - f_i)^2 \\
\text{subject to} & \;\; f_j \geq f_i + \beta_i^T (\x_j-\x_i),\; \text{for
    all $i,j$}.
\end{split}
\end{align}
See \cite{Boyd04}, Section 6.5.5.
Here $f_i$ is the estimated function value $f(\x_i)$, and the vectors
$\beta_i \in \reals^d$ represent supporting hyperplanes to the
epigraph of $f$.  Importantly, this finite dimensional quadratic program does
not have tuning parameters for smoothing the function.

\begin{figure}[t]
{\sc AC/DC Algorithm for Variable Selection in Convex Regression\hfill}
\vskip5pt
\begin{center}
\hrule
\vskip7pt
\normalsize
\begin{enumerate}
\item[] \textit{Input}:  $(\mathbf{x}_1, y_1), ..., (\mathbf{x}_n, y_n)$, regularization parameter $\lambda$.
\vskip5pt
\item[] \textit{AC Stage}:  Estimate a sparse additive convex model:
\begin{equation}
\label{eqn:scam2}
\hat{f}_1, ..., \hat{f}_p = \argmin_{f_1,...,f_p \in \mathcal{C}^1} 
   \frac{1}{n} \sum_{i=1}^n \Big(y_i - \sum_{k=1}^p f_k(x_{ik}) \Big)^2 
       + \lambda \sum_{k=1}^p \| f_k \|_\infty
\end{equation}
\vskip5pt
\item[] \textit{DC Stage}:  Estimate decoupled concave functions
 for each $k$ such that $\| \hat{f}_k \| = 0$:
\begin{equation}
\label{eqn:dc2}
\hat{g}_k = \argmin_{g_k \in \mathcal{C}^1} 
   \frac{1}{n} \sum_{i=1}^n \Big( y_i - \sum_{k'} \hat{f}_{k'}(x_{ik'}) 
    - g_k(x_{ik})\Big)^2 
      + \lambda \| g_k \|_\infty
\end{equation}
\item[] \textit{Output}: Component functions $\{\hat f_k\}$ and 
relevant variables $\hat S$ where
\begin{equation}
\hat S^c = \bigl\{k : \| \hat{f}_k \| =
0 \; \mathrm{and}\; \|\hat{g}_k \|=0\bigr\}.
\end{equation}
\end{enumerate}
\vskip3pt
\hrule
\end{center}
\vskip0pt
\caption{The AC/DC algorithm for variable selection in convex
  regression.  The AC stage fits a sparse additive convex regression
  model, using a quadratic program that imposes an group sparsity
  penalty for each component function.  The DC stage fits
  decoupled concave functions on the residuals, for each 
  component that is zeroed out in the AC stage.}
\label{fig:backfitting:algo}
\end{figure}
